### 2025-06-06 paper/conclusion.tex
Update conc, intro, abs.

This thesis bridges economic social learning theory and multi-agent reinforcement learning by introducing Partially Observable Active Markov Games (POAMGs) and the POLARIS algorithm. Our framework addresses fundamental challenges in modeling strategic adaptation under partial observability, where agents must simultaneously learn about their environment and strategically influence others' learning processes.

Our theoretical contributions establish POAMGs as a rigorous mathematical framework extending Active Markov Games to partially observable settings. The convergence analysis and policy gradient theorems provide solid foundations for the POLARIS algorithm, which successfully implements this framework through an integrated architecture combining belief processing, inference learning, and reinforcement learning modules.

Experimental validation across strategic experimentation and learning without experimentation scenarios demonstrates our framework's effectiveness. POLARIS agents converge to theoretically predicted equilibrium strategies, validating key insights about free-riding behavior and optimal experimentation intensity. Our results confirm the existence of social learning barriers while revealing coordination benefits in larger networks. Crucially, we uncover a dynamic division of labor where agents assume roles as information generators or exploiters based on signal quality rather than network position.

Future research could extend the framework to incorporate heterogeneous agent types, dynamic network structures, or mechanism design principles. Applications to contemporary challenges such as misinformation spread or algorithmic recommendation systems could demonstrate broader relevance.

This work demonstrates that integrating economic theory and computational methods yields insights neither approach could achieve alone. The POAMG framework and POLARIS algorithm provide both theoretical understanding and practical tools for analyzing complex social learning processes, establishing a foundation for future interdisciplinary research. As multi-agent systems become increasingly prevalent, such integrated approaches will be essential for understanding and shaping collective intelligence emerging from strategic interactions among learning agents.

>>>>

This thesis bridges economic social learning theory and multi-agent reinforcement learning by developing Partially Observable Active Markov Games (POAMGs) and the POLARIS algorithm. Our framework addresses fundamental challenges in modeling adaptive behavior under partial observability, demonstrating how computational approaches complement traditional economic theory while providing new insights into complex social learning dynamics.

\paragraph{Summary of Contributions}

Our work makes four interconnected contributions. First, we developed the POAMG framework, extending Active Markov Games to partially observable settings where agents reason about environmental uncertainty and evolving strategies of others. Second, we provided theoretical analysis of convergence and equilibrium properties, establishing stochastic stability conditions and deriving policy gradient theorems for average and discounted rewards. We extended the framework to continuous-time dynamics through stochastic differential equations and revealed how game-theoretic equilibrium concepts relate to active equilibrium when agents account for others' learning. Third, we introduced POLARIS, combining belief processing through Transformer models, variational inference learning with GNNs, and reinforcement learning optimization, enabling agents to develop sophisticated strategies accounting for both environmental partial observability and strategic adaptation. Fourth, we validated our framework through strategic experimentation and learning without experimentation applications, demonstrating dynamic role assignment, information generation and exploitation, and emergence of learning barriers and coordination benefits.

\paragraph{Key Findings and Implications}

Our experimental analysis reveals that dynamic role assignment is a robust organizing principle of multi-agent learning systems. Across both scenarios, agents naturally differentiate into complementary roles enhancing collective information processing. Our results show that free-riding behavior does not lead to predicted inefficiencies—performance of some agents in larger networks substantially exceeds autarky levels. The computational approach illuminates fundamental tension between theoretical and practical approaches: while economic models envision agents developing ex ante strategies optimal across all states, reinforcement learning agents converge to ex post strategies optimized for specific realized states. This reveals limitations in directly applying computational methods to test theoretical predictions while suggesting opportunities for more realistic behavioral models.

\paragraph{Methodological Contributions and Limitations}

Our two-staged analysis methodology provides systematic approaches for extracting insights from multi-agent learning dynamics despite catastrophic forgetting. The discretization approach for continuous-time economic models treats Lévy processes through Euler-Maruyama schemes and maps continuous decisions to discrete action probabilities, providing computational implementation templates while preserving essential incentive structures. We constructed observed reward functions enabling reinforcement learning without direct reward signals and developed specialized transformer loss functions for Lévy process observations. Our implementation incorporates Graph Neural Networks with temporal and spatial attention mechanisms capturing network topology and temporal dependencies. We introduce theoretically grounded off-policy asymmetric analysis methods moving beyond symmetric equilibria analysis, allowing heterogeneous agent behaviors and asymmetric strategy convergence. Despite these advances, catastrophic forgetting prevents direct measurement of predicted equilibrium strategies, requiring alternative analytical approaches, and computational complexity limits analysis to small networks and finite horizons, potentially missing asymptotic behaviors theoretical models emphasize.

\paragraph{Future Research Directions}

Several promising avenues emerge: developing continual learning techniques for social learning environments could address catastrophic forgetting through architectures maintaining separate policy components for different states; scaling to larger networks and longer horizons would enable asymptotic validation; extending to dynamic environments where underlying states change would broaden applicability; and investigating optimal information structures and network topologies could inform platform and institutional design.

\paragraph{Broader Impact and Conclusion}

This thesis demonstrates the value of interdisciplinary approaches bridging economic theory and computational methods. By preserving theoretical rigor while leveraging modern machine learning techniques, we gain insights into complex social phenomena neither approach could achieve alone. The insights have implications for financial markets, organizational learning, online social networks, and collaborative scientific endeavors. Dynamic role assignment suggests effective social learning systems naturally develop specialized roles enhancing collective information processing, providing guidance for platform design and institutional arrangements. Our POAMG framework and POLARIS algorithm contribute to multi-agent reinforcement learning by providing principled approaches for handling partial observability and strategic adaptation. The techniques may prove valuable for other applications requiring sophisticated multi-agent coordination under uncertainty. The convergence of economic theory and artificial intelligence presents exciting opportunities for advancing understanding of social learning and strategic behavior. This thesis provides theoretical foundations and computational tools for future researchers to illuminate rich dynamics of social learning in complex environments. While significant progress has been made, much work remains to fully realize this integration's potential, with the ultimate goal of better understanding how intelligent agents can learn from each other to make better decisions in an uncertain world.

### 2025-06-06 paper/intro.tex
Update conc, intro, abs.

This thesis bridges these domains by introducing a novel framework that integrates economic social learning theory with multi-agent reinforcement learning under partial observability. Our approach models social learning as a dynamic process where agents with limited information strategically adapt their behaviors while simultaneously learning about their environment and other agents' strategies. By formalizing this process through the lens of partially observable reinforcement learning, we aim to reveal new insights into how strategic adaptation influences collective outcomes and equilibrium behavior. This interdisciplinary synthesis addresses fundamental challenges that have constrained previous work in both economics and artificial intelligence, offering a more realistic and computationally tractable approach to modeling complex social learning dynamics.

@@

To address these challenges, this thesis makes several interconnected contributions. First, we develop Partially Observable Active Markov Games (POAMGs)—a novel formalism that extends the Active Markov Game framework introduced by \citet{kim2022influencing} to partially observable settings. Unlike standard reinforcement learning frameworks that treat non-stationarity as a challenge to be mitigated, POAMGs incorporate policy evolution as an integral part of the environment dynamics, allowing agents to reason about and strategically influence this evolution process. POAMGs explicitly model how agents' policies evolve over time based on observations and interactions, while accounting for the fundamental constraints imposed by partial observability. This formalism provides a mathematically rigorous foundation for analyzing complex social learning dynamics that involve repeated interactions, adaptive strategies, and incomplete information.

Second, we provide a thorough theoretical analysis of convergence and equilibrium properties in POAMGs. We establish conditions under which the joint process of states, beliefs, and policy parameters converges to a unique stochastically stable distribution, ensuring that our models have well-defined limiting behavior despite the inherent non-stationarity of multi-agent learning. We derive policy gradient theorems for average and discounted reward objectives, providing a solid foundation for algorithm development. We also extend our framework to continuous-time dynamics through stochastic differential equations, enabling a wider range of applications. Our analysis illuminates the relationship between traditional game-theoretic equilibrium concepts and the more general notion of active equilibrium that emerges in our framework, offering new insights into equilibrium selection and stability in social learning contexts.

Third, we introduce POLARIS (Partially Observable Learning with Active Reinforcement In Social environments), a practical algorithm for learning in POAMGs. POLARIS extends previous algorithms to partially observable settings through an integrated architecture with three core components: a belief processing module that tracks agents' information states using Transformer models; an inference learning module that predicts the policy evolution of other agents through variational methods; and a reinforcement learning module that optimizes policies based on average or discounted reward criteria, supporting both discrete and continuous action spaces. This integrated approach enables agents to learn sophisticated strategies that account for both the partial observability of the environment and the learning dynamics of other agents—capabilities that are essential for modeling realistic social learning processes.

Fourth, we apply our framework to two canonical social learning scenarios from economic theory. The first application focuses on strategic experimentation, building on the models of \citet{bolton1999strategic} and \citet{keller2020undiscounted}, where agents learn optimal actions through trial and error with observable rewards. The second application examines learning without experimentation, extending the frameworks of \citet{huang2024learning} and \citet{brandl2024}, where agents form beliefs primarily by observing others' actions rather than receiving direct feedback from the environment. Our analysis demonstrates how our approach can model phenomena such as free-riding on others' information production, strategic teaching, information cascades, and confounded learning under more realistic assumptions about agents' reasoning capabilities and strategic sophistication.

Through these applications, we illustrate how bridging economic theory and MARL can enhance our understanding of social learning processes. Our POAMG framework provides a more realistic model of how partially rational agents learn from each other in complex, partially observable environments. At the same time, it offers practical algorithms for computing equilibrium strategies that traditional economic approaches might find intractable. The POLARIS algorithm demonstrates how techniques from deep reinforcement learning and approximate inference can be adapted to tackle the specific challenges of social learning, offering a computationally feasible method for simulating and analyzing complex multi-agent learning dynamics.

>>>>

This thesis bridges these domains by introducing a novel framework that integrates economic social learning theory with multi-agent reinforcement learning under partial observability. Our approach models social learning as a dynamic process where agents with limited information strategically adapt their behaviors while simultaneously learning about their environment and other agents' strategies. Through the development of Partially Observable Active Markov Games (POAMGs) and the POLARIS algorithm, we demonstrate how computational approaches can complement traditional economic theory while revealing new insights into complex social learning dynamics, particularly the emergence of dynamic role assignment as a fundamental organizing principle of multi-agent learning systems.

@@

To address these challenges, this thesis makes four interconnected contributions that advance both economic theory and multi-agent reinforcement learning. First, we develop Partially Observable Active Markov Games (POAMGs)—a novel formalism that extends the Active Markov Game framework introduced by \citet{kim2022influencing} to partially observable settings. Unlike standard reinforcement learning frameworks that treat non-stationarity as a challenge to be mitigated, POAMGs incorporate policy evolution as an integral part of the environment dynamics, allowing agents to reason about and strategically influence this evolution process. POAMGs explicitly model how agents' policies evolve over time based on observations and interactions, while accounting for the fundamental constraints imposed by partial observability.

Second, we provide a thorough theoretical analysis of convergence and equilibrium properties in POAMGs. We establish conditions under which the joint process of states, beliefs, and policy parameters converges to a unique stochastically stable distribution, ensuring that our models have well-defined limiting behavior despite the inherent non-stationarity of multi-agent learning. We derive policy gradient theorems for average and discounted reward objectives and extend our framework to continuous-time dynamics through stochastic differential equations. Critically, we reveal how game-theoretic equilibrium concepts relate to active equilibrium when agents account for others' learning, illuminating the fundamental tension between theoretical ex ante strategies and the ex post strategies that emerge through reinforcement learning.

Third, we introduce POLARIS (Partially Observable Learning with Active Reinforcement In Social environments), combining belief processing through Transformer models, variational inference learning with Graph Neural Networks (GNNs), and reinforcement learning optimization. Our implementation incorporates GNNs with temporal and spatial attention mechanisms that capture network topology and temporal dependencies, enabling agents to develop sophisticated strategies accounting for both environmental partial observability and strategic adaptation. POLARIS addresses the critical challenge of catastrophic forgetting through a two-staged analysis methodology that provides systematic approaches for extracting insights from multi-agent learning dynamics despite agents' inability to maintain stable policies across all states.

Fourth, we validate our framework through strategic experimentation and learning without experimentation applications, building on the models of \citet{bolton1999strategic}, \citet{keller2020undiscounted}, \citet{huang2024learning}, and \citet{brandl2024}. Our analysis demonstrates dynamic role assignment as a robust organizing principle, where agents naturally differentiate into complementary roles enhancing collective information processing. Contrary to theoretical predictions about free-riding inefficiencies, we show that performance of some agents in larger networks substantially exceeds autarky levels, revealing opportunities for more realistic behavioral models that account for computational agents' learning dynamics.

A key methodological innovation is our discretization approach for continuous-time economic models, which treats Lévy processes through Euler-Maruyama schemes and maps continuous decisions to discrete action probabilities. This provides computational implementation templates while preserving essential incentive structures from theoretical models. We construct observed reward functions that enable reinforcement learning without direct reward signals and develop specialized transformer loss functions for Lévy process observations. Our theoretically grounded off-policy asymmetric analysis methods move beyond symmetric equilibria analysis, allowing heterogeneous agent behaviors and asymmetric strategy convergence—capabilities essential for understanding real-world social learning dynamics.

The insights gained advance theoretical knowledge and inform the design of platforms and institutions that facilitate efficient collective learning. Our finding that dynamic role assignment emerges as a natural organizing principle suggests that effective social learning systems develop specialized roles enhancing collective information processing, providing guidance for platform design and institutional arrangements in financial markets, organizational learning, online social networks, and collaborative scientific endeavors. By demonstrating the value of MARL techniques for addressing classical questions in economic theory while revealing computational challenges like catastrophic forgetting, we contribute to both fields and encourage further cross-disciplinary work leveraging complementary insights from these rich intellectual traditions.

### 2025-06-06 paper/thesis.tex
Update results.
Update conc, intro, abs.

{\Large \textbf{Social Learning via Multi Agent Reinforcement Learning}}\\[6cm]

@@

% Abstract
\begin{abstract}
\begin{adjustwidth}{2cm}{2cm}
This thesis bridges economic social learning theory and multi-agent reinforcement learning by introducing Partially Observable Active Markov Games (POAMGs)—a novel framework for modeling sophisticated learning dynamics in environments with partial observability. Traditional economic models provide valuable theoretical insights but often rely on simplified assumptions about agents' behavior or the learning environment, while reinforcement learning approaches offer computational tools that frequently lack theoretical foundations for strategic adaptation. Our framework addresses three fundamental challenges: modeling strategic adaptation in non-stationary environments, incorporating partial observability as an intrinsic feature, and capturing long-term strategic considerations in learning contexts. We develop POLARIS (Partially Observable Learning with Active Reinforcement In Social environments), a practical algorithm implementing our theoretical framework, and apply it to two canonical social learning scenarios: strategic experimentation with observable rewards and learning without experimentation through action observation. Our results validate theoretical predictions about ... while demonstrating how reinforcement learning agents can discover sophisticated strategies for information revelation and strategic influence, providing both theoretical insights and computational tools for understanding complex social learning dynamics.
\end{adjustwidth}
\end{abstract}

@@

% Page numbering starts here with Arabic numerals
\chapter{Introduction}
\pagenumbering{arabic}
\setcounter{page}{1}
\input{intro}
\chapter{Related Work}
\input{background}
\chapter{Theoretical Framework}
\input{theory}
\chapter{Results}
\input{results}
\chapter{Conclusion}
\input{conclusion}

>>>>

{\Large \textbf{Social Learning via\\[0.25cm]Multi-Agent Deep Reinforcement Learning}}\\[6cm]

@@

% Abstract
\begin{abstract}
\begin{adjustwidth}{2cm}{2cm}
This thesis bridges economic social learning theory and multi-agent reinforcement learning by developing Partially Observable Active Markov Games (POAMGs) and the POLARIS algorithm. Our framework addresses fundamental challenges in modeling adaptive behavior under partial observability, where agents must simultaneously learn about uncertain environments and evolving strategies of others. We extend Active Markov Games to partially observable settings, derive theoretical convergence results, and develop a practical algorithm combining Transformer-based belief processing, Graph Neural Networks with attention mechanisms, and reinforcement learning optimization. Applied to strategic experimentation and learning without experimentation scenarios, our approach reveals dynamic role assignment as a robust organizing principle where agents naturally differentiate into complementary roles enhancing collective information processing. We find that free-riding behavior does not lead to uniform inefficiencies—for some agents, performance in larger networks exceeds autarky levels. Our two-staged analysis methodology addresses computational challenges like catastrophic forgetting while providing systematic approaches for extracting insights from multi-agent learning dynamics. The framework contributes theoretical foundations to multi-agent reinforcement learning and computational tools to economic theory, demonstrating how interdisciplinary approaches can illuminate complex social learning phenomena neither field could address alone.
\end{adjustwidth}
\end{abstract}

@@

% Page numbering starts here with Arabic numerals
\chapter{Introduction}
\pagenumbering{arabic}
\setcounter{page}{1}
\input{intro}
\chapter{Related Work}
\input{background}
\chapter{Reinforcement Learning Framework}
\input{theory}
\chapter{Evaluation}
\input{results}
\chapter{Conclusion}
\input{conclusion}
